Distributed multi-objective evolutionary optimization using island-based selective operator application

### 2019-02-13 hpmoon-asoco.tex
Added reference for issue #8 and some suggestions by @JJ, including the automatic->automatically typo

%%%NEW AND SHORTED
Parallel island-model co-evolutionary algorithms are well-known methods, suitable for dealing with large multi-objective optimization problems. In these algorithms, each island modifies a fragment of the chromosome that encodes a possible solution to the problem. In this paper, we want to demonstrate that automatically setting the size of the overlapping fragments depending on the number of islands obtains better results than using the same size. We  have compared this method to other parallel evolutionary techniques considering a different number of islands, chromosome sizes and benchmarks. The analysis of the obtained experimental results, by using different metrics, shows that our approach can provide statistically significant improvements with respect to the base algorithm in high-dimensional, un-decomposable, multi-objective problems. This opens a very challenging line to automatically adapt the overlapping sizes in this kind of algorithms.

@@

\subsection{Proposed methods}
In this paper we analyze several SLA algorithms we have
implemented by using a parallel multi-objective baseline algorithm
as a basis for comparisons. Our proposed methods, using different overlapping schemes, %methods or method?
                                %Which methods? - JJ Fergu: clarifying
are based on NSGA-II, as almost all the previous works discussed
before
\cite{Dorronsoro13superlinear,Durillo08masterslave,Hiroyasu07discussion,Deb03distributed,Xiao03specialized,Wang09parallel,Martens13asynchronous}. % This is SoA stuff - JJ FERGU: Yes, I added it here againg because a reviewer wasn't very attentive
Therefore, we have used a basic distributed NSGA-II algorithm without
overlapping sections as the baseline (B).

@@

As in the baseline (B), an individual is migrated to another random
island after a fixed number of generations. In the new island, this
individual will be considered as one of the others in the island,
crossed and mutated in the same way, depending on the island
identifier (from 1 to $P$). Note that, unlike the other works such as the ones
described by Talbi et al. \citep{Talbi08Parallel}, all the islands deal
with complete chromosomes for fitness calculation, so our approach can
deal with decomposable and no decomposable problems. %Don't understand what you say
                                %here. What you are saying is that our
                                %approach can't deal with decomposable
                                %problems, actually. - JJ FERGU: you are right, fixed

@@

\subsubsection{\textcolor{blue}{Automatic} SLA with overlapping islands (A)} 
As in the previous method, the sections to address by the operators are overlapped, but
instead of using one extra section on each side of the $p$ section ($p+1$
and $p-1$), it uses $c$ fragments on each side ($p+c$ and $p-c$),
being $c$ a value depending on the number of islands.
% That's not adaptive. It's simply variable - JJ FERGU: it depends on the number of islands. What term do you suggest?
% FERGU: changed to Automatic
As a first
approach to automatically calculate this value, the results shown in
\citep{Garcia16hpmoon} have been used as a base to obtain $c$. In that
work, the overlapped method ($c=1$) obtained better results when the
number of islands \textcolor{blue}{was} bigger than 8. On the contrary, when the number of
islands is small, no overlapping is necessary. So, we have used this
knowledge to propose the formula $c=round(0.2*P-1)/2$ to calculate the
extra sections to overlap. Therefore, when $P=8$ then $c=0$
(equivalent to D), when $P=16$ then $c=1$ (equivalent to O), and so
on. Dividing individuals by blocks of equal size, instead \textcolor{blue}{of} a percentage, allow the future integration of our method with other recombination mechanisms, such as the ones presented in \cite{Dorronsoro13superlinear} or \cite{Kimovski15Parallel}.
Also, as mentioned, this is only a possible formula to reach above-described behaviour for $c$, and future work will study new methods or alternative formulas to determine it, for example using Taguchi Method \cite{Keratmatpour2018Taguchi}.  Figure \ref{fig:adaptive} explains the $A$ method, assuming that $c=2$, for example.

@@

We have chosen those metrics not only because they have been used
extensively, especially in some of the papers presented in Section \ref{sec:soa}
% Still section 2? Please create a reference and use it - JJ Pablo: done
\cite{Dorronsoro13superlinear,Durillo08masterslave,Hiroyasu07discussion,Wang09parallel,Martens13asynchronous},
but also because they cover different quality criteria. Mathematical
formulation of these metrics can be found in
\cite{Dorronsoro13superlinear}.

@@

This migration
  is performed asynchronously, as it has the advantage of being a
  non-blocking operation \cite{TalbiUnified2018}. The individual
  chosen to migrate is selected from the subpopulation using Binary
  Tournament from a randomly selected couple, as proposed in previous papers
  \cite{xiong2003parallel,Xiao03specialized}. Whenever an island
  starts the migration step, the migrated individual is sent to
  another randomly selected island. This makes it possible a
  dissemination of the modified parts of the individuals among the
  whole set of islands, thus avoiding a specific synchronization step
  to compose solutions obtained by different islands, as occurs other
  works \cite{Dorronsoro13superlinear}.

 The whole parameter set is summarized in Table \ref{tab:parameters}. 
The main parameters in the proposed system were the number of islands and the size
   of the problem, as pointed out by other authors 
   \cite{CaoZLL17,Dorronsoro13superlinear,DorronsoroPSO2018,Martens13asynchronous,Durillo08masterslave}. This is why those have been the focus of this particular
   article. Although all the parameters indicated above have an
   influence on the general result of the MOEA, their general effect
   is well understood and, although there will be an interplay between
 them and the number of islands and problem size, this is left mainly
 as future work.

@@

\end{tabular}
}
\caption{Average quality metrics obtained after 30 runs per
  configuration, for the 4 methods compared: baseline (B), disjoint
  (D), overlapped (O) and \textcolor{blue}{automatic} overlapped
  (A), % No sería mejor "automatically overlapped" or "automatic
       % overlap"? This does not sound good - JJ
  using a chromosome length of 512 dimensions. Acronyms next to values
  indicate that there is not significant difference with respect to
  that method for that value. Best values are marked in
  bold. Equiv-$X$ implies that value is the same as executing $X$,
  since both configurations would be the same.}  
\label{tab:results512}
\end{table*}

@@

\end{tabular}
}
\caption{Average quality metrics obtained after 30 runs per configuration, for the 4 methods compared: baseline (B), disjoint (D), overlapped (O) and \textcolor{blue}{automatic} overlapped (A), using a chromosome length of 2048 dimensions. Acronyms next to values indicate that there is not significant difference with respect to that method for that value. Best values are marked in bold. Equiv-$X$ implies that value is the same as executing $X$, since both configurations would be the same.}
\label{tab:results2048}
\end{table*}

@@

\caption{Average number of generations and  solutions per front, obtained after 30 runs per configuration, for the 4 methods compared: baseline (B), disjoint (D), overlapped (O) and \textcolor{blue}{automatic} overlapped (A), using a chromosome length of 512 dimensions. Acronyms next to values indicate that there is not significant difference with respect to that method for that value. Equiv-$X$ implies that value is the same as executing $X$, since both configurations would be the same.}

@@

\caption{Average number of generations and solutions per front after 30 runs per configuration, for methods baseline (B), disjoint (D), overlapped (O) and \textcolor{blue}{automatic} overlapped (A), using a chromosome length of 2048 dimensions. Acronyms next to values indicate that there is not significant difference with respect to that method for that value. Equiv-$X$ implies that value is the same as executing $X$, since both configurations would be the same.}

@@

%%%%%%FERGU: WARNING, THIS SECTION IS STILL SHIT
%ZDT1 
After establishing the performance of the new method and in order to
have some insight on how they work, let's check the effect of the different chromosome sizes and number of
islands. Figures \ref{fig:plot_zdt1_512} to \ref{fig:plot_zdt6_2048} show the
final PFs of one arbitrary sample run of each combination;. The first
one, ZDT1, has a convex Pareto-optimal front. Comparing results from
$L=512$ (Figure \ref{fig:plot_zdt1_512}) and $L=2048$ (Figure
\ref{fig:plot_zdt1_2048}) it is apparent that the quality of the
generated PFs depends on the chromosome size. Specifically, all
configurations  with
$L=512$  seem to have a worse Spread and HV, on the other hand, the
SLA methods behave worse than $B$, as it was also shown on average in
Table \ref{tab:results512}. Nevertheless, if we focus on the plots
obtained with $L=2048$ (Figure \ref{fig:plot_zdt1_2048}) this
difference is less clear. Figure \ref{fig:plot_zdt1_2048} shows that
solutions obtained by the SLA methods show higher dispersion than the
baseline as $P$ grows. However, with $A$, the area covered by the PF
is still larger.

@@

Obtained results also show that when increasing the number of islands, the
automatic overlapping method ($A$) significantly improves the results with
respect to the disjoint and overlapped methods. However, this is only
apparent with larger chromosome sizes. Therefore, the length of the
chromosome is also a key factor to take into account in automatic overlapping
methods, and not only the number of islands. Studying this factor with
more types of problems, and new configurations of population size and
chromosome lengths could be addressed in the future. Since our method divides the individuals into blocks of divisible size, aggregation methods after migration, such as the ones used in \cite{Kimovski15Parallel} or \cite{Dorronsoro13superlinear} can be also tested. But as we are dealing with complete individuals in each island, the possibility of using a real number, such as 0.3 (30\% overlapping) instead an integer value opens future experimental setups.
For example, comparing 
different ways to calculate $c$ using linear, logarithmic or exponential functions
depending on population size, number of islands, or other values, or using some statistical parameter setting method, such as Taguchi's \cite{Keratmatpour2018Taguchi}. Moreover, a diversity analysis of the individuals
in each island during the algorithm run could be carried out to understand its influence
on the results.

>>>>

%%%NEW AND SHORTED
Parallel island-model co-evolutionary algorithms are well-known methods, suitable for dealing with large multi-objective optimization problems. In these algorithms, each island modifies a fragment of the chromosome that encodes a possible solution to the problem. In this paper, we want to demonstrate that automatically setting the size of the overlapping fragments depending on the number of islands obtains better results than using the same size. We  have compared this method to other parallel evolutionary techniques considering a different number of islands, chromosome sizes and benchmarks. The analysis of the obtained experimental results, by using different metrics, shows that our approach can provide statistically significant improvements with respect to the base algorithm in high-dimensional, un-decomposable, multi-objective problems. This opens a very promising line to automatically adapt the overlapping sizes in this kind of algorithms.

@@

\subsection{Proposed methods}
In this paper we analyze several SLA algorithms we have
implemented by using a parallel multi-objective baseline algorithm
as a basis for comparisons. Our proposed methods, using different overlapping schemes, 
are based on NSGA-II, as almost all the previous works discussed
before
\cite{Dorronsoro13superlinear,Durillo08masterslave,Hiroyasu07discussion,Deb03distributed,Xiao03specialized,Wang09parallel,Martens13asynchronous}. 
Therefore, we have used a basic distributed NSGA-II algorithm without
overlapping sections as the baseline (B).

@@

As in the baseline (B), an individual is migrated to another random
island after a fixed number of generations. In the new island, this
individual will be considered as one of the others in the island,
crossed and mutated in the same way, depending on the island
identifier (from 1 to $P$). Note that, unlike the other works such as the ones
described by Talbi et al. \citep{Talbi08Parallel}, all the islands deal
with complete chromosomes for fitness calculation, so our approach can
deal with decomposable and no decomposable problems.

@@

\subsubsection{\textcolor{blue}{Automatic} SLA with overlapping islands (A)} 
As in the previous method, the sections to address by the operators are overlapped, but
instead of using one extra section on each side of the $p$ section ($p+1$
and $p-1$), it uses $c$ fragments on each side ($p+c$ and $p-c$),
being $c$ a value depending on the number of islands.

As a first
approach to automatically calculate this value, the results shown in
\citep{Garcia16hpmoon} have been used as a base to obtain $c$. In that
work, the overlapped method ($c=1$) obtained better results when the
number of islands \textcolor{blue}{was} bigger than 8. On the contrary, when the number of
islands is small, no overlapping is necessary. So, we have used this
knowledge to propose the formula $c=round(0.2*P-1)/2$ to calculate the
extra sections to overlap. Therefore, when $P=8$ then $c=0$
(equivalent to D), when $P=16$ then $c=1$ (equivalent to O), and so
on. Dividing individuals by blocks of equal size, instead \textcolor{blue}{of} a percentage, allow the future integration of our method with other recombination mechanisms, such as the ones presented in \cite{Dorronsoro13superlinear} or \cite{Kimovski15Parallel}.
Also, as mentioned, this is only a possible formula to reach above-described behaviour for $c$, and future work will study new methods or alternative formulas to determine it, for example using Taguchi Method \cite{Keratmatpour2018Taguchi}.  Figure \ref{fig:adaptive} explains the $A$ method, assuming that $c=2$, for example.

@@

We have chosen those metrics not only because they have been used
extensively, especially in some of the papers presented in Section \ref{sec:soa}
\cite{Dorronsoro13superlinear,Durillo08masterslave,Hiroyasu07discussion,Wang09parallel,Martens13asynchronous},
but also because they cover different quality criteria. Mathematical
formulation of these metrics can be found in
\cite{Dorronsoro13superlinear}.

@@

This migration
  is performed asynchronously, as it has the advantage of being a
  non-blocking operation \cite{TalbiUnified2018}. The individual
  chosen to migrate is selected from the subpopulation using Binary
  Tournament from a randomly selected couple, as proposed in previous papers
  \cite{xiong2003parallel,Xiao03specialized}. Whenever an island
  starts the migration step, the migrated individual is sent to
  another randomly selected island. This makes it possible a
  dissemination of the modified parts of the individuals among the
  whole set of islands, thus avoiding a specific synchronization step
  to compose solutions obtained by different islands, as occurs other
  works \cite{Dorronsoro13superlinear}.\textcolor{blue}{The main parameters studied in the proposed system are limited to the number of islands and the size of the problem. This decision has been motivated because all previous works described in the state of the art also limit their experimental setting to the same parameters \cite{CaoZLL17,Dorronsoro13superlinear,DorronsoroPSO2018,Martens13asynchronous,Durillo08masterslave}. Although all the parameters indicated above have an influence on the general result of the MOEA, their general effect is well understood and, although there will be an interplay between them and the number of islands and problem size, this is left mainly as future work.}

@@

\end{tabular}
}
\caption{Average quality metrics obtained after 30 runs per
  configuration, for the 4 methods compared: baseline (B), disjoint
  (D), overlapped (O) and \textcolor{blue}{automatically} overlapped
  (A), % No sería mejor "automatically overlapped" or "automatic
       % overlap"? This does not sound good - JJ FERGU: Done
  using a chromosome length of 512 dimensions. Acronyms next to values
  indicate that there is not significant difference with respect to
  that method for that value. Best values are marked in
  bold. Equiv-$X$ implies that value is the same as executing $X$,
  since both configurations would be the same.}  
\label{tab:results512}
\end{table*}

@@

\end{tabular}
}
\caption{Average quality metrics obtained after 30 runs per configuration, for the 4 methods compared: baseline (B), disjoint (D), overlapped (O) and \textcolor{blue}{automatically} overlapped (A), using a chromosome length of 2048 dimensions. Acronyms next to values indicate that there is not significant difference with respect to that method for that value. Best values are marked in bold. Equiv-$X$ implies that value is the same as executing $X$, since both configurations would be the same.}
\label{tab:results2048}
\end{table*}

@@

\caption{Average number of generations and  solutions per front, obtained after 30 runs per configuration, for the 4 methods compared: baseline (B), disjoint (D), overlapped (O) and \textcolor{blue}{automatically} overlapped (A), using a chromosome length of 512 dimensions. Acronyms next to values indicate that there is not significant difference with respect to that method for that value. Equiv-$X$ implies that value is the same as executing $X$, since both configurations would be the same.}

@@

\caption{Average number of generations and solutions per front after 30 runs per configuration, for methods baseline (B), disjoint (D), overlapped (O) and \textcolor{blue}{automatically} overlapped (A), using a chromosome length of 2048 dimensions. Acronyms next to values indicate that there is not significant difference with respect to that method for that value. Equiv-$X$ implies that value is the same as executing $X$, since both configurations would be the same.}

@@

%%%%%%FERGU: WARNING, THIS SECTION IS STILL SHIT
%ZDT1 
\textcolor{blue}{After establishing the performance of the new method and in order to
have some insight on how they work, let's check the effect of the different chromosome sizes and number of
islands. Plotting the solution sets allows to observe its quality, distribution and shape, specially if they are limited to 2 or 3 objectives, as in this paper \cite{Li17read}. } Figures \ref{fig:plot_zdt1_512} to \ref{fig:plot_zdt6_2048} show the
final PFs of one arbitrary sample run of each combination;. The first
one, ZDT1, has a convex Pareto-optimal front. Comparing results from
$L=512$ (Figure \ref{fig:plot_zdt1_512}) and $L=2048$ (Figure
\ref{fig:plot_zdt1_2048}) it is apparent that the quality of the
generated PFs depends on the chromosome size. Specifically, all
configurations  with
$L=512$  seem to have a worse Spread and HV, on the other hand, the
SLA methods behave worse than $B$, as it was also shown on average in
Table \ref{tab:results512}. Nevertheless, if we focus on the plots
obtained with $L=2048$ (Figure \ref{fig:plot_zdt1_2048}) this
difference is less clear. Figure \ref{fig:plot_zdt1_2048} shows that
solutions obtained by the SLA methods show higher dispersion than the
baseline as $P$ grows. However, with $A$, the area covered by the PF
is still larger.

@@

Obtained results also show that when increasing the number of islands, the
automatic overlapping method ($A$) significantly improves the results with
respect to the disjoint and overlapped methods. However, this is only
apparent with larger chromosome sizes. Therefore, the length of the
chromosome is also a key factor to take into account in automatic overlapping
methods, and not only the number of islands. Studying this factor with
more types of problems, and new configurations of population size and
chromosome lengths could be addressed in the future. Since our method divides the individuals into blocks of divisible size, aggregation methods after migration, such as the ones used in \cite{Kimovski15Parallel} or \cite{Dorronsoro13superlinear} can be also tested. But as we are dealing with complete individuals in each island, the possibility of using a real number, such as 0.3 (30\% overlapping) instead an integer value opens future experimental setups.
For example, comparing 
different ways to calculate $c$ using linear, logarithmic or exponential functions
depending on population size, number of islands, or other values, or using some statistical parameter setting method, such as Taguchi's \cite{Keratmatpour2018Taguchi}. Moreover, a diversity analysis of the individuals
in each island during the algorithm run could be carried out to understand its influence
on the results. \textcolor{blue}{Parallel coordinates will be used when dealing with problems with more than 2 objectives \cite{Li17read}}.
